---
title: "Assignment"
author: "Thi Thu Huong Nguyen"
date: "5/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1\
**a.**

```{r surg}
surg <- read.csv("~/R/surg.dat", sep="")
surg1= subset(surg, select=-gender)
pairs(surg1, panel=panel.smooth)
```

-   Relationship between the response and predictors\
    As we can see in the plot, there are possible relationships among survival time and other predictors. While the relationships between Survival with blood, prognosis, enzyme and liver are moderately or slightly positive, its between survival and age is negative. The strongest positive relationship is between survival and liver.

-   Relationship among predictors\
    There are some noticeable positive relationships between liver with other predictors such as blood, prognosis and enzyme. The relationships between age and other predictors can be seen to be weak from the scatter plot.

-   We need to remove *gender* variables because it's categorical variable. In order to compute correlation matrix, we need numerical variable.

**b.**

```{r}
cor(surg1)
```


By computing correlation matrix, we can test the correlation strength among the response and predictors. Overall, there are mostly weak correlation, raging from -0.2 to 0.3. However, there are still some moderate correlation. Moderate correlation: survival with enzyme, prognosis and liver; liver with blood and enzyme.

**c.**\
Since the researchers try to use many predictor variables to explain the survival response, multiple regression model is suitable.

```{r}
surg1_lm = lm(survival ~ blood + prognosis + enzyme + liver + age, data=surg1)
```

*Mathematics multiple regression model*\
$$
Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4 + \beta_5X_5 + \epsilon;  \epsilon \sim Norm(0,\sigma^2)
$$
Y: survival  
$\beta_0$: \text{Intercept term}  
$\beta_1, \beta_2,..., \beta_5$ :\text{ Partial regression coefficients}  
$X_1 : blood ; X_2 : prognosis ; X_3 : enzyme ; X_4 : liver ; X_5 : age$  
$\epsilon$ : \text{variation a way from regression line.}  

\text{Test hypothesis:}  
$H_0$ : $\beta_1 = \beta_2 = \beta_3 = \beta_4 = \beta_5$  
$H_1$ : not all $\beta_i$ = 0

```{r}
anova(surg1_lm)
```
**_ANOVA table for the overall multiple regression model_**

|   Source    |  df  |   Sum of Sq   |    Mean Sq    |    F   |
|:------------|-----:|--------------:|--------------:|-------:|
|  Regression |  5   |  5816714      | 1163342.8     | 21.874 |
|  Residual   |  48  |  2552807      | 53183         |        |
|------------ |----- |-------------- |-------------- |--------|
|   Total     |  53  |  8369521      |               |        |

* Test statistic: $F_{obs}$ =   $\frac{RegressionM.S}{ResidualM.S}$ = $\frac{1163342.8}{53183}$ = 21.874
* Null distribution: $F_{g,n-g-1}$ : $F_{5,48}$
* P-Value: P($F_{5,48} \ge 21.874$) = 2.386226e-11 < 0.05
* Reject at the 5% level.
* There is a significant linear relationship between survival and at least one of the five predictor variables.

**d.**
To find the best multiple regression model, we can apply stepwise backward estimation. 
```{r}
summary(surg1_lm)$coefficients
```
Drop liver variable since it has the largest p-value (4.375949e-01)
```{r}
summary(lm(survival ~ blood + prognosis + enzyme + age, data= surg1))$coefficients
```

Age is insignificant after dropping liver. Other variables are significant. 
```{r}
summary(lm(survival ~ blood + prognosis + enzyme, data= surg1))$coefficients
```

Blood, prognosis and enzyme are significant. This is the best multiple regression model. 

**e.** 
Validate the best multiple regression model
```{r}
best_model_lm = lm(survival ~ blood + prognosis + enzyme, data= surg1)

par(mfrow = c(1, 2))
plot(best_model_lm, which = 1:2)

```
Comment: 

* The QQ plot follows a positive linear, however, there are some observations such as #5, #28 and #54 may cause future problems.

* Residual vs fitted plot shows curvature, where the non-linear relationship cannot be explained by the model. 

Therefore, the model is not validate and not appropriate to use the multiple regression model to explain the survival time. 


**f.**
```{r}
new_survival= log(surg$survival)
new_surg1_lm = lm(new_survival ~ blood + prognosis + enzyme + liver+age, data= surg1)
```

```{r}
summary(new_surg1_lm)$coefficients
```
Remove liver since it has the largest p-value.
```{r}
summary(lm(new_survival ~ blood + prognosis + enzyme + age, data= surg1))$coefficients
```
After dropping liver, all variables are significant except for age. Age should be dropped.

```{r}
summary(lm(new_survival ~ blood + prognosis + enzyme, data= surg1))$coefficients
```
All variables now are significant. This is the best model. 

**g.**
Validate the new best model:

```{r}
new_best_model_lm = lm(new_survival ~ blood + prognosis + enzyme, data= surg1)

par(mfrow = c(1, 2))
plot(new_best_model_lm, which = 1:2)

```
Comment: In the residuals vs Fitted plot, all the data points are distributed more evenly than the previous best model. Although there are still curves, it is more prone to straight line than the previous one. The normal QQ plot also shows a stronger linear line, which shows an improvement in the new model. Overall, the model with log(survival) is superior to the one with survival as the response. 

### Question 2\
**a.**
```{r}
kml <- read.csv("~/R/kml.dat", sep="")
table(kml[, c("driver","car")])
```
It is a balanced design because the number of driver and car is the same.

**b.**
```{r}
boxplot(kmL ~ car + driver, data = kml)
```
With each car and driver, the differences vary in the range from 10 to 16 kmL. Moreover, means of each observation slightly vary while the spread are moderately similar. 

```{r}
attach(kml)
interaction.plot(driver, car,kmL, col = 1:5, fixed=TRUE)
```
This plot is able to show clearly the difference among drivers and cars on Kml. While driver A and C seem to share the same fuel efficiency from 10 to 13, driver B shows the greatest efficiency, from 13 to 16. Cars 2, 4, 5, 1 and 3 show the order of efficiency from greatest to lowest, respectively. 

**c.**\
Because the question requires to test the interaction between factors, it is relevant to apply Two-way ANOVA.

$$ 
Y_{ijk} = \mu + \alpha_{i} + \beta_{j} + \gamma_{ij} + \epsilon_{ijk}, \epsilon_{ijk} \sim {\sf N}(0,\sigma^2) 
$$ 

$\mu$ = overall population mean.   
$\alpha_{i}$ : base effect of $i^{th}$ level of Factor driver: i = 1,2,...,a.   
$\beta_{j}$ : base effect of $j^{th}$ level of Factor car: j = 1,2,...,b.   
$\gamma_{ij}$ : effect of the combined effect of the $i^{th}$, $j^{th}$ \text{combination of the two factors.}   
$\varepsilon_{ijk}$ : \text{unexplained variation for each replicated observation,} $\varepsilon \sim {\sf N}(0,\sigma^2)$ .   

* Hypothesis about the interaction term:   
$H_0$ : $\gamma_{ij}$ = 0 for all i,j ; $H_A$ : not all $\gamma_{ij}$ = 0

```{r}
kml_lm = lm(kmL ~ driver * car, data = kml)
anova(kml_lm)
```
Interaction is not significant (0.3715>0.05). Therefore, we remove the interaction term. 

```{r}
par(mfrow=c(1,3))
plot(kml_lm, which = 1:2)
hist(kml_lm$residuals)
```
The residuals vs fitted plot show evenly distributed data points. The QQ plot indicates a straight line and the histogram shows a slightly right-skewed bar graph, but it can be considered to be normally distributed. 

**d.**\
The fuel efficiency reaches its greatest with car two and driver B, while reaches it lowest with car three and driver A. As we can see from the model and the graphs, both car and driver can affect the fuel efficiency. 